A modal logical multiverse for commutative algebra and combinatorics

### 2022-11-04 slides-paris2022.tex
More graphics

@@

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{5.95cm}\includegraphics[width=\paperwidth]{fr1-lighter}\end{minipage}}
\begin{frame}{Three proofs}
  \begin{center}\bignumber{1}\end{center}
  \vspace*{-1.3em}
  \begin{varblock}{\textwidth}{}
    \justifying
    \textbf{Theorem.}
    For every number~$n$, there is a prime larger than~$n$.
  \end{varblock}
  \vspace*{-0.4em}
  {\emph{Proof.} Any prime factor of~$n! + 1$ will do. \qed}
  \pause
  \medskip

  \begin{center}\bignumber{2}\end{center}
  \vspace*{-1.3em}
  \begin{varblock}{\textwidth}{}
    \justifying
    \textbf{Theorem.}
    Every infinite sequence~$\alpha : \NN \to \NN$ is \emph{good}
    in that there are numbers~$i < j$ such that~$\alpha(i) \leq \alpha(j)$.
  \end{varblock}
  \vspace*{-0.4em}
  {\emph{Proof.} By~\badbox{\textsc{lem}}, there is a minimal value~$\alpha(i)$.
  Set~$j \defeq i + 1$.\qed}
  \pause
  \medskip

  \begin{center}\bignumber{3}\end{center}
  \vspace*{-1.3em}
  \begin{varblock}{\textwidth}{}
    \justifying
    \textbf{Theorem.}
    Let~$M$ be a surjective matrix with more rows than columns over a
    commutative ring~$A$.
    Then~$1 = 0$ in~$A$.
  \end{varblock}
  \vspace*{-0.4em}
  {\justifying\emph{Proof.}
  \only<3>{\bad{Assume not.} Then there is~a \bad{maximal ideal} $\mmm$.
  The matrix is surjective over~$A/\mmm$, contradicting
  basic linear algebra.}\only<4>{Write~$M = \left(\begin{smallmatrix}x\\y\end{smallmatrix}\right)$. By surjectivity,
  have~$u, v$ with
  $
    u \left(\begin{smallmatrix}x\\y\end{smallmatrix}\right) = \left(\begin{smallmatrix}1\\0\end{smallmatrix}\right)
  $ and $
    v \left(\begin{smallmatrix}x\\y\end{smallmatrix}\right) = \left(\begin{smallmatrix}0\\1\end{smallmatrix}\right).
  $
  Hence
  $
    1 = (vy) (ux) = (uy) (vx) = 0
  $.}
  \qed\par}
\end{frame}}
% XXX: chart of ideals of ℤ

@@

  \begin{columns}[c]
    \begin{column}{0.25\textwidth}
      \includegraphics[width=\textwidth,valign=c]{multiverse}
    \end{column}
    \begin{column}{0.75\textwidth}
      There is a \hil{rich multiverse} of \\
      flavors of parametrized mathematics.
    \end{column}
  \end{columns}
\end{frame}

>>>>

\input{images/primes.tex}

@@

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{5.95cm}\includegraphics[width=\paperwidth]{fr1-lighter}\end{minipage}}
\begin{frame}{Three proofs}
  \begin{columns}[c]
    \begin{column}{0.15\textwidth}
      \scalebox{0.23}{\sieve{6}{6}}
    \end{column}
    \begin{column}{0.85\textwidth}
      \begin{varblock}{\textwidth}{}
        \justifying
        \textbf{Thm.}
        For every~$n$, there is a prime larger than~$n$.
      \end{varblock}
      \vspace*{-0.4em}
      \emph{Proof.} Any prime factor of~$n! + 1$ will do. \qed
    \end{column}
  \end{columns}
  \pause

  \begin{columns}[c]
    \begin{column}{0.15\textwidth}
      \includegraphics[width=\textwidth,valign=t]{monad}
    \end{column}
    \begin{column}{0.85\textwidth}
      \begin{varblock}{\textwidth}{}
        \justifying
        \textbf{Thm.}
        Every infinite sequence~$\alpha : \NN \to \NN$ is \emph{good}
        in that there are numbers~$i < j$ such that~$\alpha(i) \leq \alpha(j)$.
      \end{varblock}
      \vspace*{-0.4em}
      \mbox{\emph{Proof.} There is a \bad{minimal value}~$\alpha(i)$.
      Set~$j \defeq i{+}1$.\qed}
    \end{column}
  \end{columns}
  \pause

  \begin{columns}[c]
    \begin{column}{0.15\textwidth}
      \includegraphics[width=\textwidth,valign=t]{calabi-yau}
    \end{column}
    \begin{column}{0.85\textwidth}
      \begin{varblock}{\textwidth}{}
        \justifying
        \textbf{Thm.}
        Let~$M$ be a surjective matrix with more rows than columns over a
        ring~$A$. Then~$1 = 0$ in~$A$.
      \end{varblock}
      \vspace*{-0.4em}
      \justifying\emph{Proof.}
      \only<3>{\bad{Assume not.} Then there is~a \bad{maximal ideal} $\mmm$.
      The matrix is surjective over~$A/\mmm$. Since~$A/\mmm$ is a field, this
      is a contradiction to basic linear algebra. \\[1.3em]\ }\only<4>{Write~$M =
      \left(\begin{smallmatrix}x\\y\end{smallmatrix}\right)$. By surjectivity,
      have~$u, v$ with
      \vspace*{-0.6em}
      \[
        u \left(\begin{smallmatrix}x\\y\end{smallmatrix}\right) = \left(\begin{smallmatrix}1\\0\end{smallmatrix}\right)
        \quad\text{and}\quad
        v \left(\begin{smallmatrix}x\\y\end{smallmatrix}\right) = \left(\begin{smallmatrix}0\\1\end{smallmatrix}\right).
      \]

      \vspace*{-0.6em}
      Hence
      $
        1 = (vy) (ux) = (uy) (vx) = 0
      $.}
      \qed
    \end{column}
  \end{columns}
\end{frame}}
% XXX: chart of ideals of ℤ

% \begin{document}

@@

  \begin{columns}[c]
    \begin{column}{0.25\textwidth}
      \includegraphics[width=\textwidth,valign=c]{multiverse}
    \end{column}
    \begin{column}{0.75\textwidth}
      There is a \hil{rich modal multiverse} of \\
      flavors of parametrized mathematics.
    \end{column}
  \end{columns}

  \visible<3->{\begin{tikzpicture}[overlay]
    \node[anchor=south east,inner sep=0] (image) at (10.8,0.3) {
      \reflectbox{\includegraphics[width=1em]{butterfly}}
    };
  \end{tikzpicture}}
\end{frame}
